---
title: "Data Analysis Assignment 02 - Resume Data"
format: html
editor: visual
jupyter: ir
---

```{r}
#Importing Libraries
library(dplyr)
library(caret)
#Reading the data
library(openintro)
data("resume")

#Storing it in a dataframe called 'base_data'
base_data <- resume
```

```{r}
head(resume)
print(nrow(base_data))
print(ncol(base_data))
```

```{r}
glimpse(base_data)
```

```{r}

```

# Model Documentation

Research Question: How do race and gender influence job application callback rates?

### 1. Overview

The resume data in the OpenIntro Library is a dataset of Resumes that were used to apply for job profiles, and whether or not they recieved a callback. The resume dataset contains the following fields -

-   *Job Details* - These include details such as City, Industry, Job Title, Private/Non Profit, required education, and required skills
-   *Applicant Details* - Details about the applicant, such as Gender, Race, years of education, college degree, skills, and years of experience
-   *Resume Details* - Details about the resume, such Email available, Resume Quality
-   *Callback* - whether the applicant received a call back for this job posting for their resume (1 or 0) - this will be the *dependent variable*

The dataset will be used to train a logistic regression model to predict the probability of receiving an interview invite, given the gender and socioeconomic class of the applicant.

### 2. Data Cleaning

```{r}
resume$gender_factors <- factor(resume$gender, levels = c("m", "f"))
resume$resume_quality_factors <- factor(resume$resume_quality, levels = c("low", "high"))
resume$race_factors <- factor(resume$race, levels = c("black", "white"))
resume$job_equal_opp_employer <- as.factor(resume$job_equal_opp_employer)
resume$job_fed_contractor <- as.factor(resume$job_fed_contractor)
resume$job_req_any <- as.factor(resume$job_req_any)
resume$job_req_communication <- as.factor(resume$job_req_communication)
resume$job_req_education <- as.factor(resume$job_req_education)
resume$job_req_min_experience <- as.factor(resume$job_req_min_experience)
resume$job_req_computer <- as.factor(resume$job_req_computer)
resume$job_req_organization <- as.factor(resume$job_req_organization)
resume$honors <- as.factor(resume$honors)
resume$worked_during_school <- as.factor(resume$worked_during_school)
resume$computer_skills <- as.factor(resume$computer_skills)
resume$special_skills <- as.factor(resume$special_skills)
resume$volunteer <- as.factor(resume$volunteer)
resume$military <- as.factor(resume$military)
resume$employment_holes <- as.factor(resume$employment_holes)
resume$has_email_address <- as.factor(resume$has_email_address)
```

### 3. Modeling

### 4. Results

### 5. Future Work
